SiO Molecular Jets around young stars - A numerical perspective

### 2013-07-27 PAPER/ms_final_bv.tex
Added Introduction and plots

\newcommand{\rhoi}{{\rho}_{\rm i}}

@@

\section{Introduction} 
Interaction of jet with ambient medium , Class 0 sources, momentum and
energy injection, High mass and low mass outflows \\

Observations of Outflows in CO and SiO in mm, H2 in IR, Optical jets
HH objects, syncrotron emission \\

Molecules from jet driven or wind driven, Interesting features like
EHV, low velocity component etc, list interesting outflows from Class 0 \\

Attempts of simulations either only dynamics with analytic approaches
to compare observational features, or only chemistry to estimate
abundances. The reality lies in the middle!\\

Combine dynamics of MHD jet with molecular chemistry and comparing
with observations via post-processing simulations with line radiative
transfer codes. \\

Paper Layout \\

@@

\section{Results}
\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/jetplt1_1010.png}
 \caption{Jet Volume Density for different cooling modes with
   $\eta$ = 10 and $\beta$ = 10.}
\label{grid}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/molcjet_510.png}
 \caption{Fraction of hydrogen species in the run with molecular
   cooling having $\eta$ = 5 and $\beta$ = 10.}
\label{grid}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/linejets.png}
 \caption{SiO emission from different line transitions for the run with molecular
   cooling having $\eta$ = 3 and $\beta$ = 10. }
\label{grid}
\end{figure*}

\begin{table*}
\centering
\caption{Summary from parameter runs}
\begin{tabular}{l | l | l | l | l | l }
\hline
Run & Cooling Mode & $\eta$ & $\beta$ & Peak Intensity [K] & FWHM at Bow
Shock \\
\hline\hline
adi1010 & Nil (Adiabatic) & 10 & 10 &&\\
pow1010 & Power law & 10 & 10 &&\\
atm1010 & Atomic & 10 & 10 &&\\
atm101 & Atomic & 10 & 1 &&\\
atm210 & Atomic & 2 & 10 &&\\
tab1010 & Tabulated & 10 & 10 &&\\
tab210 & Tabulated & 2 & 10&&\\
mol1010 & Molecular & 10 & 10&&\\
mol510 & Molecular & 5 & 10&&\\
\hline
\end{tabular}
\label{tab:result1}

@@

@@

\begin{figure*}
 \includegraphics[width=168mm]{\figpath/tabulated_mom0maps_Trans.png}
 \caption{The integrated intensity of 4 SiO transitions in the tabulated cooling model in the $\eta$=2 $\beta$=10 model.}
\label{tabulatedTrans}
\end{figure*}

\begin{figure}
 \includegraphics[width=84mm]{\figpath/tab_trans_spectra.png}
 \caption{The Integrated emission and spectra at he bow shock for 2 transitions of the tabulated cooling model for $\eta$=2 $\beta$=10 (spectra taken from y=50, x=128).}
\label{grid}
\end{figure}

@@

Spectra and PV Digrams\\
Line - withs $\sim$ 15-25 km\,s$^{-1}$. Wedge shaped PV diagrams
a signature feature of jet driven molecular outflows.

Comparison with Observed results \\
Different line transition trace different components of SiO
emission. The low level transitions trace the region where the jet
interacts with molecular medium. The clumpy nature in its emission is
from the turbulent nature of its interaction. The higher transitions
trace a much more collimated region along with the terminal shock. 
As seen by Chandler \& Richer 2001 and Hirano 2006.

>>>>

\newcommand{\rhoi}{{\rho}_{\rm i}}
\newcommand{\refer}{{\color{orange}REFERENCE}}

@@

\section{Introduction}
Jets are one of the first manifestations of star
formation in dense molecular cores. They are ubiquitous in both
massive and low star forming regions. These supersonic flows
perpendicular to the underlying accertion disk plays a vital role in
removing excess angular momentum and thereby aiding in the accretion. 
For low mass stars, they are believed to be launched by magneto-centrifugally forces and
further collimated by magnetic hoop stress
\citep[][]{Blandford:1982p892, Konigl:2000p607}. However, in case of
high mass stars, radiative forces also contribute to flow dynamics
during the later evolutionary stages \cite{Vaidya:2011p8992}. 
Typically, these jets are few parsec size long and can be divided into
three length-scale domains viz. source and disk scales (1-10$^{2}\,$AU),
envelope scales ($10^{2} - 10^{5}$\,AU) and parent cloud scales
($10^{5} - 10^{6}$\,AU) \refer. Among them, the envelope scales are
the ones where rich chemical evolution occurs as the jet interacts
with the molecular medium. In this region, the jet propagates into a
relatively static medium inducing shocks that are interesting from the
physical and chemical point of view. In addition to shocks, molecular
material from the surrounding is entrained and accelerated to high
velocities giving rise to molecular outflows.
%

Bipolar molecular outflows from low and high mass stars
have been studied in details over the past 
decade \citep[see reviews by][]{Bachiller:1996p4692,Arce:2007p798,
Tafalla:2011p14051}. Advancement in millimeter interferometers have
allowed to observe these outflows with high spatial resolution of
few arc seconds. A large number of studies related to young
outflows are done using standard outflow and shock tracer
like CO and SiO respectively. In addition to these molecular tracers,
shocks from these outflows are detected in molecular hydrogen using
infra-red telescopes. Based on these observational studies, various
empirical properties for these outflows have been discovered. For
example, the CO outflows from single dish studies are seen to be
highly collimated for both low and high mass stars
\citep[for e.g.,][]{Gueth:1999p4683, Beuther:2002p3574} they are
referred to as {\em{molecular jets}}. These outflows also exhibit a
mass-velocity relation with a power law form $ dM(v)/dv \propto
v^{-\gamma}$, where values of $\gamma$ range from 1 to 3 \cite{Downes:2003p9946}. Episodic
knots belived to be caused by variable accretion events are a common property
of young molecular outflows (for e.g. in L1577:
\citealt{Gueth:1998p14058}, in HH300: \citealt{Arce:2001p14064}). These knots show their signatures
as {\em{wedges}} in the position-velocity (PV) diagrams \cite{Arce:2001p14065}. Also, most
commonly observed in these outflows are signatures of rotation and
precision (for e.g., in DG Tau: \citealt{Bacciotti:2002p2084}).
%

Even with myriad of empirical evidences, the exact nature of SiO and
CO outflow is not clear. For a complete understanding it is imperative
to compliment these observations with theoretical models. Many models
based on hydro-simulations and steady state shock calculations
were proposed to explain the observational signatures of molecular
outflows. Among them the two main models are that of wide-angled wind
driven \cite{Shu:1991p14071} and jet driven outflow
\cite{Canto:1991p14123}. 
The most popular among them is the jet driven model as wind
driven molecular outflows not only fail to match observed PV digarams \cite{Cabrit:1992p14098}
but also tend to sweep large quantitity of material at the extermities
of the lobes \cite{Masson:1992p14101}. While the jet driven models could sucessfully
derive the global outflow shapes and mass velocity relations of CO
outflows \citep{Raga:1993p9948, Masson:1993p9661} . There have also been some attempts to combine these two
models into one \cite{Shang:2006p14268} to explain the global observational features. 
However, most of these dynamical models do not
account for shock chemsitry. Instead, shock chemistry is studied
independently without dynamics using plane parallel shock
models \citep{Schilke:1997p14140,Flower:2003p11236} . The magneto-hydrodynamic (MHD) calculations by \cite{Glassgold:1991p13703} suggested that molecules
like SiO and CO could as form within the jet. Similar conclusions of
molecules surving in steady state disk winds have also been shown
\cite{Panoglou:2012p10039}. Very limited simulations have modelled the outflow
dynamics including molecular chemistry but in absence of
magnetic fields \citep{Raga:1995p12965, Smith:2003p9985}.

%
In the present work, our goal is to take a step further in modelling of
jet driven molecular outflows. The present model aims to consistently derive observed emission
properties of molecular outflows, specifically various SiO line
tranistions, by combining axisymmetric MHD simulations of
radiative jet propogation. In particular, we evolve the jet dynamical
quantities in conjunction with different non-equilibrium cooling
prescriptions of varying complexities. The most complex is that of
molecular cooling along with H$_{2}$ chemistry. This prescription
allows us to track the formation and destruction of 
HI, HII and H2 along with the flow dynamics. 
The final state of the jet obtained for each cooling model is then
post-processed using a non-LTE radiative line transfer code
to obtain emission maps, spectra and PV digram. These emission maps
are further processed using ALMA-CASA pipeline to obtain synthetic
images of molecular outflows.
%

In the next three section we describe our numerical setup, cooling
presecriptions and radiative transfer line code respectively. In
Sect. 5, we will present results from the parameter survey and the
discussions along with predicted ALMA maps will be presented in
Section 6 and 7, followed by conclusions.

@@

\section{Parameter Survey}
\begin{table*}
\centering
\caption{Summary from parameter runs}
\begin{tabular}{l | l | l | l | l | l }
\hline
Run & Cooling Mode & $\eta$ & $\beta$ & Peak Intensity [K] & FWHM at Bow
Shock \\
\hline\hline
adi1010 & Nil (Adiabatic) & 10 & 10 &&\\
pow1010 & Power law & 10 & 10 &&\\
atm1010 & Atomic & 10 & 10 &&\\
atm101 & Atomic & 10 & 1 &&\\
atm210 & Atomic & 2 & 10 &&\\
tab1010 & Tabulated & 10 & 10 &&\\
tab210 & Tabulated & 2 & 10&&\\
mol1010 & Molecular & 10 & 10&&\\
mol510 & Molecular & 5 & 10&&\\
\hline
\end{tabular}
\label{tab:result1}

@@

\section{Results}
\subsection{Comparison of cooling prescription}
figure nos. 2 -- Similuated images, emission maps for eta = 10.
\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/jetplt1_1010.png}
 \caption{Jet Volume Density for different cooling modes with
   $\eta$ = 10 and $\beta$ = 10.}
\label{grid}
\end{figure*}

@@

\subsection{Molecular cooling and H$_2$ Chemistry}
figure nos. 2 --> Explain molecular cooling in more details. With the
mechanism of formation and destruction of HI, H2 and HII using
figures. 
\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/molcjet_310.png}
 \caption{Fraction of hydrogen species in the run with molecular
   cooling having $\eta$ = 3 and $\beta$ = 10.}
\label{grid}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/molcjet_310_molform.png}
 \caption{Dependence of hydrogen fractions on the temperature at two
   points in the flow, viz. the interface of the knot with molecular
   medium and at the bow shock}
\label{grid}
\end{figure*}

\subsection{Effects of changing $\eta$ and $\beta$}
figure nos. 2 -- Simulated images describing lengths, collimation and
emission for lower eta more stronger. 
\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/jetplt2_diffetabeta.png}
 \caption{Jet Volume Density for atomic cooling with $\eta$ = 10 but $\beta$ = 1 and
   10 and that with tabulated cooling with $\beta$ = 10 but $\eta$ = 2
   and 10.}
\label{grid}
\end{figure*}

\subsection{SiO Abundance profile}
figure nos. 2 -- Comparison of three profiles for SiO Abundance,
Contour maps for two transitions for top hat profile.

\begin{figure}
 \includegraphics[width=1.\columnwidth]{\figpath/contourmapsSiOTopHat.png}
 \caption{Contours of SiO 2-$>$1 and 8-$>$7 emission obtained using the
   SiO Top Hat profile.}
\label{grid}
\end{figure}

\subsection{Spectra and PV diagrams}
figure nos. 1 -- one transition 3 different angles, Comparison of
spectra and line ratios

@@

Effects of varing abundance profile ... Crucial ingredient in
modeling molecular bullets \\

Spectra and PV Digrams\\
Line - withs $\sim$ 15-25 km\,s$^{-1}$. Wedge shaped PV diagrams
a signature feature of jet driven molecular outflows. Counterparts in observations.

Comparison with Observed results \\
Different line transition trace different components of SiO
emission. The low level transitions trace the region where the jet
interacts with molecular medium. The clumpy nature in its emission is
from the turbulent nature of its interaction. The higher transitions
trace a much more collimated region along with the terminal shock. 
As seen by Chandler \& Richer 2001 and Hirano 2006.
\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/linejets.png}
 \caption{SiO emission from different line transitions for the run with molecular
   cooling having $\eta$ = 3 and $\beta$ = 10. }
\label{grid}
\end{figure*}

Advantage of non-LTE modeling over LVG Cabrit 2007, Lee 2008. 
Line ratios $\sim$ 1 can be explained well. \\

EHV component ?? \\
